Formal Synthesis of Digital Controllers for Observable Continuous Plants with Transient Performance Specifications

### 2018-01-31 TACAS2018/paper.tex
line breaks

Modern implementations of embedded control systems have proliferated with
the availability of low-cost devices that can perform highly non-trivial
control tasks, with significant impact in numerous application areas such as
life sciences, environmental control and robotics~\cite{astrom1997computer,Franklin15}.  Correct synthesis is non-trivial, even in cases with
restricted dynamics.  We examine the case of Linear Time Invariant (LTI)
models, for which the synthesis of controllers is well understood.  However,
the use of digital control architectures adds new challenges caused by
artifacts specific to digital control, such as the effects of
finite-precision arithmetic (referred to here as Finite Word Length or FWL effects), time discretization and quantization errors
introduced by A/D and D/A conversion.

@@

In this paper we build on the ideas presented in~\cite{abate2017automated}
and develop a new technique that addresses most of these limitations. 
By~combining state-of-the-art synthesis and verification engines we
automatically generate digital observer controllers for a given continuous
plant model that are correct by construction.  %% This approach delivers a high
%% degree of automation, promises to reduce the cost and time of development of
%% digital control dramatically, and requires considerably less expertise than
%% a-posteriori verification.
Specifically, we synthesize stable,
software-implemented embedded controllers along with a model of a physical
plant (observer).  Due to the complexity of such closed-loop systems, we focus
on linear models with known configurations, and perform
parametric synthesis of controllers meeting a number of properties that
define their desired behavior.

@@

We use a CounterExample-Guided Inductive Synthesis (CEGIS)
\cite{jha-icse10,DBLP:conf/pldi/Solar-LezamaRBE05} approach,
which we enhance with abstraction refinement and optimization
resulting in the novel CEGIS-OR (CEGIS with Optimization and
Refinement).  This design uses an iterative process where each
iteration performs inductive generalization based on counterexamples
provided by a verification oracle.  Essentially, the inductive
generalization uses information about a limited number of inputs to
make claims in the form of candidate solutions about all the possible
inputs.  By using abstraction refinement, CEGIS-OR allows us to handle
complex systems which are initially represented by coarse abstractions
that only get refined on demand.  Consequently, solutions that do not
require a high level of precision are obtained fast.  Since the use
of coarse abstractions may result in low-quality candidate solutions
(i.e. candidates that fail the specification on a very large number of
inputs), after each inductive generalization phase we use optimization
techniques to improve the produced candidates.  In particular, our
optimization procedure aims at reducing the size of the reach space of
the closed loop system, while maintaining other dynamical properties
of interest (safety, stability, and response time).

@@

There are two different problems that arise from FWL semantics.  The first
is the error in the dynamics caused by the inability to represent the exact
state of the physical system, while the second relates to rounding and
saturation errors during computation.  In~\cite{fialho1994stability}, a
stability measure based on the error of the digital dynamics ensures that
the deviation introduced by FWL does not make the digital system unstable. 
A~more recent approach~\cite{DBLP:journals/automatica/WuLCC09} uses
$\mu$-calculus to directly model the digital controller so that the selected
parameters are stable by design.  The analyses
in~\cite{DBLP:conf/hybrid/RouxJG15,DBLP:conf/hybrid/WangGRJF16} rely on an
invariant computation on the discrete system dynamics using Semi-Definite
Programming (SDP).  While the former uses bounded-input and bounded-output
(BIBO) properties to determine stability, the latter uses Lyapunov-based
quadratic invariants.  In both cases, the SDP solver uses floating-point
arithmetic and soundness is checked by bounding the error.  An alternative
is~\cite{park2016scalable}, where the verification of given control code is
performed against a known model by extracting an LTI model of the code by
symbolic execution: to account for rounding errors, an upper bound is
introduced in the verification phase.  The work in
\cite{picasso2003stabilization} introduces invariant sets as a mechanism to
bound the quantization error effect on stabilization as an invariant set
that always converges toward the controllable set.  Similarly,
\cite{liberzon2003hybrid} evaluates the quantization error dynamics and
bounds its trajectory to a known region over a finite time period.  This
technique works for both linear and non-linear systems.

@@

We consider models of physical plants expressed as ordinary differential
equations (ODEs), which we assume to be controllable \cite{Astrom08,astrom1997computer}:
%
\begin{align}
\label{eq:ode1}
\dot{x}(t) = \mat{A}\vec{x}(t)+ \mat{B} \vec{u}(t), \quad \vec{x} \in \mathbb{R}^p, \vec{u} \in \mathbb{R}^q, \mat{A} \in \mathbb{R}^{p \times p}, \mat{B} \in \mathbb{R}^{p \times q},
\end{align}
%
where $t \in \mathbb R_0^+$ denotes continuous time, 
$\mat{A}$ and $\mat{B}$ are matrices that fully characterize the dynamics of the continuous plant, 
and $x(0)$ is the initial state.

We are interested in control systems with an output $\vec{y}$ that is a linear combination of its states and inputs\footnote{Notice that since we can only measure $\vec{y}$, the state-space $\vec{x}$ is not fully observable.}.
%, which may restrict the observability of the state-space from the output space.
\begin{equation}
\label{eq:ode2}  
\vec{y}(t)=\mat{C}\vec{x}(t)+\mat{D}\vec{u}(t), \vec{y} \in \mathbb{R}^o, \mat{C} \in \mathbb{R}^{o \times p}, \mat{D} \in \mathbb{R}^{o \times q}. 
\end{equation}
%
Equations~\eqref{eq:ode1}-\eqref{eq:ode2} are soundly discretized in
time~\cite{middleton1990digital,van1978computing} at constant intervals, whose 
duration we refer to as the sample time $T_s$, 
into the difference equation 
\begin{align}
\label{eq:plant}
\left\{
\begin{array}{rl}
\vec{x}_{k+1} =& \mat{A}_d \vec{x}_k+ \mat{B}_d \vec{u}_k\\
\vec{y}_{k} =& \mat{C} \vec{x}_k + \mat{D} \vec{u}_k, 
\end{array}
\right.
\text{ where }
\left\{
\begin{array}{l}
\mat{A}_d=e^{\mat{A}T_s}\\
\mat{B}_d = \int_{t = 0}^{T_s} e^{\mat{A} t} dt\ \mat{B}
\end{array}
\right.,
\end{align} 
%
%where $\mat{A}_d=e^\mat{A}$, $\mat{B}_d = \int_{t = 0}^{T_s} e^{\mat{A} t} dt\ \mat{B}$,
$k \in \mathbb N$ is a discrete counter and $\vec{x}_{0}=\vec{x}(0)$ is the initial state. 
In this work we raise the standard assumption that $\mat{D}=\mat{0}$ \cite{Astrom08}.  
%Also, since $\vec{y}$ can be directly derived from $\vec{x}$, we will omit discussing it in the following unless strictly necessary.

@@

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{State observers}\label{sec:observer}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
In many models of real-world control systems, 
the internal state $x$ cannot be directly measured,  
but is rather inferred from its effect on the measured output $\vec{y}$ using an observer~\cite{Astrom08}.  
Essentially, a state observer is a new dynamical model (with variable denoted as $\hat{\vec{x}}$) 
that provides an estimate of the internal state $x$ of the given model, 
from measurements of its inputs and outputs.

@@

In Figure \ref{fig:observersystem} we have introduced sources of noise that are due to quantization effects.  
These effects are caused by imprecise numerical representations of real numbers in the digital domain of the controller and play an important role in the synthesis of sound controllers. 
The model has two sources of error due to Finite Word Length (FWL) representations. 
The first is the numerical error introduced by the fixed-point numbers employed to model the plant, 
i.e., to represent the plant dynamics $A_d$, $B_d$, $C_d$, $D_d$, $x_k$ and $y_k$. 
The second is the quantization error introduced by the digital controller, 
which performs operations over fixed-point numbers.  
In this section we outline the notation for the finite representation of numbers, 
and briefly describe the introduced errors.  
A formal discussion is in Appendix~\ref{sec:numeric_rep}.

Let $\mathbb{R}_{\langle M,F \rangle}$ be a fixed point domain with $M$ bits
representing the binary word length of the representation and $F$ bits
representing the pre-scaling $2^{-F}$ (\emph{i.e} the number of bits after
the decimal place).  We note that a negative $F$ denotes an integer
multiplication that adds $-F$ trailing zeros in the number represented by
$M$. 
We will use the domain $\mathbb{R}_{\langle C \rangle} \text{ where } \langle C \rangle = \langle M_c,F_c \rangle$
to represent the fixed point precision of the controller, and 
$\mathbb{R}_{\langle P \rangle} \supseteq \mathbb{R}_{\langle C \rangle}$
to represent the domain of a digital model of the dynamics of a real plant when
evaluated by a verification tool.  Elements in this domain are marked with an
underscore~${\langle C \rangle}$ (see Figure \ref{fig:observersystem}).
Thus, any mathematical operation within the digital controller is executed in the
domain $\mathbb{R}_{\langle C \rangle}$, and all other calculations
in our model will be carried out in the domain $\mathbb{R}_{\langle P \rangle}$.
Given ${\langle M,F \rangle}$, 
we denote as $\mathcal{F}_{\langle M,F \rangle} : \mathbb{R} \rightarrow \mathbb{R}_{\langle M,F \rangle}$ the corresponding truncation function. 
We further note that the results of mathematical operations in these domains are also truncated. This is referred to as roundoff errors.

@@

Since we are interested in synthesizing controllers with observers, 
in this section we will look at checking stability, safety and the response time in the presence of an observer.
%\footnote{A full description of the theory related to observer design is in Appendix \ref{sec:control_theory}.}

@@

We assume that the domain of the observer model is discretized, 
i.e., the domain for both the controller and the observer is $\mathbb{R}_{\langle C \rangle}$.
This means that we have matrices 
${\mat{A}_d}_{\langle C
\rangle},{\mat{B}_d}_{\langle C \rangle},\mat{C}_{\langle C \rangle},
\mat{K}_{\langle C \rangle}, \mat{L}_{\langle C \rangle} \in
\mathbb{R}_{\langle C \rangle}^*$, 
where $*$ represents the appropriate dimension for each matrix.
New problems arise due to the FWL truncation of these matrices. 
These are reflected in the canonical form of the resulting model (see $\mat{A}_e$ and $\mat{B}_e$),
which is expressed by:
%
{\setlength{\belowdisplayskip}{2pt}
\begin{align}
\left [\begin{array}{c}\vec{x}'\\ \vec{x}_e' \end{array}\right]_{k+1}
=& \left [\begin{array}{cc}\mat{A}_{c}+\mat{B}_{e}&\mat{B}_{o}\\ \mat{A}_{e}&\mat{A}_{o}\end{array}\right]
\left [\begin{array}{c}\vec{x}'\\ \vec{x}_e' \end{array}\right]_k
+\mat{B}'\left [\begin{array}{c} \vec{r}_k + \nu_2+\nu_2'\\ \nu_1+\nu_1'+\nu_t\end{array}\right], %\nonumber\\
%\vec{y}_k= &\mat{C}\mat{T}_c\vec{x}_k', 
% + \mat{D}_{d}\vec{u}_k
\label{eq:observer_LTI_FWL_cf}
\end{align}
}
{\setlength{\abovedisplayskip}{2pt}
\setlength{\belowdisplayskip}{2pt}
\begin{align*}
\mat{A}_c=&\left[\begin{array}{cc}
\begin{array}{ccc}k_{p-1}-a_{p-1}&\cdots&k_1-a_1\end{array}&k_0-a_0\\
\mat{I}&\begin{array}{c}0\\ \vdots \\ 0\end{array}
\end{array}\right]&
\mat{A}_o=&\left[\begin{array}{cc}
\begin{array}{c}l_{p-1}-a_{p-1}\\ \vdots\\ l_1-a_1\end{array}&\mat{I}\\
l_0-a_0&\begin{array}{ccc}0&\cdots&0\end{array}
\end{array}\right], 
\end{align*}
}
where $\vec{x}'$ is the state of the model in canonical form and $\vec{x}_e'$ is the error of the observer evaluated in the transformed domain; 
further, $\mat{A}_{c}$ the closed-loop matrix of the controller in canonical form, 
$\mat{A}_{o}$ the closed-loop matrix of the observer in canonical form, 
$\mat{B}_{o}$ the a feedback gain matrix, 
 $\mat{B}'$ the noise sensitivity matrix, 
and $\mat{A}_{e}$ and $\mat{B}_{e}$ are the error caused by the FWL truncation, seen after the transformation to canonical form.
%and we have introduced the following terms: 
%{\setlength{\abovedisplayskip}{2pt}
%\setlength{\belowdisplayskip}{2pt}
%\begin{align*}
%\mat{A}_{c}=&\mat{T}_c\left(\mat{A}_d-\mat{B}_d\mat{K}_{\langle C \rangle}\right)\mat{T}_c^{-1}&
%\mat{A}_{o}=&\mat{T}_o\left({\mat{A}_d}_{\langle C \rangle}-\mat{L}_{\langle C \rangle}\mat{C}_{\langle C \rangle}\right)\mat{T}_o^{-1}\\
%\mat{B}_{o}=&\mat{T}_c{\mat{B}_d}_{\langle C \rangle}\mat{K}_{\langle C \rangle}\mat{T}_o^{-1}&
%\mat{B}_{e}=&\mat{T}_c(\mat{B}_d-{\mat{B}_d}_{\langle C \rangle})\mat{K}_{\langle C \rangle}\mat{T}_o^{-1}\\
%\mat{A}_{e}=&\mat{T}_o\left(\mat{A}_d-{\mat{A}_d}_{\langle C \rangle}-\mat{L}_{\langle C \rangle}\mat{C}_e\right)\mat{T}_o^{-1}&
%\mat{C}_{e}=&\mat{C}-\mat{C}_{\langle C \rangle}
%\end{align*}
%}
%{\setlength{\abovedisplayskip}{0pt}
%\begin{align*}
%\mat{A}'=&\left [\begin{array}{cc}\mat{A}_{c}+\mat{B}_{e}&\mat{B}_{o}\\ \mat{A}_{e}&\mat{A}_{o}\end{array}\right]&
%\mat{B}'=&\left[\begin{array}{cc}\mat{T}_c{\mat{B}_d}_{\langle C \rangle}&\mat{0}\\\mat{0}&-\mat{T}_o\mat{L}_{\langle C \rangle}\end{array}\right]\\
%\eta_1=&\nu_1+\nu_1'+\nu_t& \eta_2=&\nu_2+\nu_2',
%\end{align*}
%}
Moreover, $\nu_1$ and $\nu_2$ are the quantization noises introduced by the ADC and the DAC, $\nu_t$ is the noise introduced by the delay, and $\nu_1'$ and $\nu_2'$ are quantization noises caused by round-offs in the FWL operations.

In a real domain $\mathbb{R}$, the characteristic polynomial
of the closed loop dynamics is defined by the characteristic polynomials of~$\mat{A}_c$
and~$\mat{A}_o$.  
\begin{align*}
P(z)=P_K(z)P_L(z)=\left(z^p+\sum_{i=0}^{p-1}{(a_i-k_i)z^{i}}\right) \left(z^p+\sum_{i=0}^{p-1}{(a_i-l_i)z^{i}}\right)
\end{align*}
Jury's criteria uses these coefficients to ensure that no root of the polynomial has a magnitude greater than one.
%\note{we should perhaps provide a qualitative hint on what are the conditions of JC? I'm afraid leaving it to the polynomials here isn't quite complete or clear.}
The introduction of the FWL errors causes this to no longer
be true.  We must therefore find a suitable solution to maintain this independence of polynomials and use the given coefficients by bounding their error w.r.t. the true coefficients of the closed loop matrix.

@@

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Safety of a digitally observed system}\label{sec:cof_safety}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
A safety specification leads to a requirement on the states of
the model. Consequently, the controller must ensure that the state
never violates the requirement.  In~this work, the safety property is
expressed as:
%
\begin{align}
\phi_\mathit{safety}^n: = & \forall k \in [0, ... , n], \mat{G}\vec{x}_k \leq \vec{h} \text{ and } \underline{y} \leq y_k \leq \overline{y}\nonumber\\
\phi_\mathit{safety}: = & \phi_\mathit{safety}^\infty
\label{eq:safetyspec}
\end{align}
%
where $\mat{G}$ and $\vec{h}$ define a set of inequalities that describe 
a safe set, within the state-space $\mathbb{R}^p$.
%
Note that FWL effects are not explicitly captured here
as they are implicit from Equation \eqref{eq:observer_LTI_FWL_cf}.

@@

%-------------------------------
\subsection{Transient response} 
\label{sec:transientspecification}
%-------------------------------
In many cases the response time of the system, i.e. how fast it converges to the set point, is a requirement of great importance. 
This is given as a ``settling time.''

\begin{theorem}
Given a set of eigenvalues for a discrete time model of a $p^{th}$ order stable system $\{\lambda_i =e^{\omega_i} \mid i \in [1, ...\,,p], |\lambda_i| \leq 1 \}$, the settling time of the system is upper-bounded by
\begin{equation}
t_s<-log_{\overline{\lambda}}({P_s}), \text{ where }  \overline{\lambda} = \max(|\lambda_i|) \text{ and } \overline{\lambda}<1,\ \forall i \in [1, ...\,,p]
\label{eq:set_time}
\end{equation}
where $0\leq P_s \leq 1$ is the desired portion of the target within which the output must remain after $t_s$.  For the proof see Appendix~\ref{sec:proofs}
\end{theorem}

@@

%The CEGIS-OR loop for this approach is described next.
Throughout the CEGIS-OR loop we make use of a specification $\phi$ that deals with the abstraction of the model dynamics.
Note that since the model abstraction gets refined at each iteration, 
$\phi$ will also be adjusted accordingly.
Initially, $\phi$ is simply given as the defined 
$\phi := \phi_{stability} \wedge \phi_{input} \wedge \phi_{init}$.

@@

In embedded software programs interacting and/or simulating physical plants it
is of great importance to understand the effects of the chosen numeric
representation for the continuous state-space. 
These representations are almost invariably approximations to the original which

>>>>

Modern implementations of embedded control systems have proliferated with
the availability of low-cost devices that can perform highly non-trivial
control tasks, with significant impact in numerous application areas such as
life sciences, environmental control and robotics~\cite{astrom1997computer,
Franklin15}.  Correct synthesis is non-trivial, even in cases with
restricted dynamics.  We examine the case of Linear Time Invariant (LTI)
models, for which the synthesis of controllers is well understood.  However,
the use of digital control architectures adds new challenges caused by
artifacts specific to digital control, such as the effects of
finite-precision arithmetic (referred to here as Finite Word Length or FWL
effects), time discretization and quantization errors introduced by A/D and
D/A conversion.

@@

In this paper we build on the ideas presented in~\cite{abate2017automated}
and develop a new technique that addresses most of these limitations. 
By~combining state-of-the-art synthesis and verification engines we
automatically generate digital observer controllers for a given continuous
plant model that are correct by construction.  %% This approach delivers a high
%% degree of automation, promises to reduce the cost and time of development of
%% digital control dramatically, and requires considerably less expertise than
%% a-posteriori verification.
%
Specifically, we synthesize stable, software-implemented embedded
controllers along with a model of a physical plant (observer).  Due to the
complexity of such closed-loop systems, we focus on linear models with known
configurations, and perform parametric synthesis of controllers meeting a
number of properties that define their desired behavior.

@@

We use a CounterExample-Guided Inductive Synthesis (CEGIS)~\cite{jha-icse10,
DBLP:conf/pldi/Solar-LezamaRBE05} approach, which we enhance with
abstraction refinement and optimization resulting in the novel CEGIS-OR
(CEGIS with Optimization and Refinement).  This design uses an iterative
process where each iteration performs inductive generalization based on
counterexamples provided by a verification oracle.  Essentially, the
inductive generalization uses information about a limited number of inputs
to make claims in the form of candidate solutions about all the possible
inputs.  By using abstraction refinement, CEGIS-OR allows us to handle
complex systems which are initially represented by coarse abstractions that
only get refined on demand.  Consequently, solutions that do not require a
high level of precision are obtained fast.  Since the use of coarse
abstractions may result in low-quality candidate solutions (i.e.  candidates
that fail the specification on a very large number of inputs), after each
inductive generalization phase we use optimization techniques to improve the
produced candidates.  In particular, our optimization procedure aims at
reducing the size of the reach space of the closed loop system, while
maintaining other dynamical properties of interest (safety, stability, and
response time).

@@

There are two different problems that arise from FWL semantics.  The first
is the error in the dynamics caused by the inability to represent the exact
state of the physical system, while the second relates to rounding and
saturation errors during computation.  In~\cite{fialho1994stability}, a
stability measure based on the error of the digital dynamics ensures that
the deviation introduced by FWL does not make the digital system unstable. 
A~more recent approach~\cite{DBLP:journals/automatica/WuLCC09} uses
$\mu$-calculus to directly model the digital controller so that the selected
parameters are stable by design.  The analyses
in~\cite{DBLP:conf/hybrid/RouxJG15, DBLP:conf/hybrid/WangGRJF16} rely on an
invariant computation on the discrete system dynamics using Semi-Definite
Programming (SDP).  While the former uses bounded-input and bounded-output
(BIBO) properties to determine stability, the latter uses Lyapunov-based
quadratic invariants.  In both cases, the SDP solver uses floating-point
arithmetic and soundness is checked by bounding the error.  An alternative
is~\cite{park2016scalable}, where the verification of given control code is
performed against a known model by extracting an LTI model of the code by
symbolic execution: to account for rounding errors, an upper bound is
introduced in the verification phase.  The work in
\cite{picasso2003stabilization} introduces invariant sets as a mechanism to
bound the quantization error effect on stabilization as an invariant set
that always converges toward the controllable set.  Similarly,
\cite{liberzon2003hybrid} evaluates the quantization error dynamics and
bounds its trajectory to a known region over a finite time period.  This
technique works for both linear and non-linear systems.

@@

We consider models of physical plants expressed as ordinary differential
equations (ODEs), which we assume to be controllable~\cite{Astrom08,
astrom1997computer}:
%
\begin{align}
\label{eq:ode1}
\dot{x}(t) = \mat{A}\vec{x}(t)+ \mat{B} \vec{u}(t), \quad \vec{x} \in \mathbb{R}^p, \vec{u} \in \mathbb{R}^q, \mat{A} \in \mathbb{R}^{p \times p}, \mat{B} \in \mathbb{R}^{p \times q},
\end{align}
%
where $t \in \mathbb R_0^+$ denotes continuous time, $\mat{A}$ and $\mat{B}$
are matrices that fully characterize the dynamics of the continuous plant,
and $x(0)$ is the initial state.

We are interested in control systems with an output $\vec{y}$ that is a linear combination of its states and inputs\footnote{Notice that since we can only measure $\vec{y}$, the state-space $\vec{x}$ is not fully observable.}.
%, which may restrict the observability of the state-space from the output space.
\begin{equation}
\label{eq:ode2}  
\vec{y}(t)=\mat{C}\vec{x}(t)+\mat{D}\vec{u}(t), \vec{y} \in \mathbb{R}^o, \mat{C} \in \mathbb{R}^{o \times p}, \mat{D} \in \mathbb{R}^{o \times q}. 
\end{equation}
%
Equations~\eqref{eq:ode1}-\eqref{eq:ode2} are soundly discretized in
time~\cite{middleton1990digital,van1978computing} at constant intervals,
whose duration we refer to as the sample time $T_s$, into the difference
equation
%
\begin{align}
\label{eq:plant}
\left\{
\begin{array}{rl}
\vec{x}_{k+1} =& \mat{A}_d \vec{x}_k+ \mat{B}_d \vec{u}_k\\
\vec{y}_{k} =& \mat{C} \vec{x}_k + \mat{D} \vec{u}_k, 
\end{array}
\right.
\text{ where }
\left\{
\begin{array}{l}
\mat{A}_d=e^{\mat{A}T_s}\\
\mat{B}_d = \int_{t = 0}^{T_s} e^{\mat{A} t} dt\ \mat{B}
\end{array}
\right.,
\end{align} 
%
%where $\mat{A}_d=e^\mat{A}$, $\mat{B}_d = \int_{t = 0}^{T_s} e^{\mat{A} t} dt\ \mat{B}$,
$k \in \mathbb N$ is a discrete counter and $\vec{x}_{0}=\vec{x}(0)$ is the initial state. 
In this work we raise the standard assumption that $\mat{D}=\mat{0}$ \cite{Astrom08}.  
%Also, since $\vec{y}$ can be directly derived from $\vec{x}$, we will omit discussing it in the following unless strictly necessary.

@@

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{State observers}\label{sec:observer}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++

In many models of real-world control systems, the internal state $x$ cannot
be directly measured, but is rather inferred from its effect on the measured
output $\vec{y}$ using an observer~\cite{Astrom08}.  Essentially, a state
observer is a new dynamical model (with variable denoted as $\hat{\vec{x}}$)
that provides an estimate of the internal state $x$ of the given model, from
measurements of its inputs and outputs.

@@

In Figure~\ref{fig:observersystem} we have introduced sources of noise that
are due to quantization effects.  These effects are caused by imprecise
numerical representations of real numbers in the digital domain of the
controller and play an important role in the synthesis of sound controllers. 
The model has two sources of error due to Finite Word Length (FWL)
representations.  The first is the numerical error introduced by the
fixed-point numbers employed to model the plant, i.e., to represent the
plant dynamics $A_d$, $B_d$, $C_d$, $D_d$, $x_k$ and $y_k$.  The second is
the quantization error introduced by the digital controller, which performs
operations over fixed-point numbers.  In this section we outline the
notation for the finite representation of numbers, and briefly describe the
introduced errors.  A formal discussion is in
Appendix~\ref{sec:numeric_rep}.

Let $\mathbb{R}_{\langle M,F \rangle}$ be a fixed point domain with $M$ bits
representing the binary word length of the representation and $F$ bits
representing the pre-scaling $2^{-F}$ (\emph{i.e} the number of bits after
the decimal place).  We note that a negative $F$ denotes an integer
multiplication that adds $-F$ trailing zeros in the number represented
by~$M$.  We will use the domain $\mathbb{R}_{\langle C \rangle}
\text{ where } \langle C \rangle = \langle M_c,F_c \rangle$ to represent the fixed point
precision of the controller, and $\mathbb{R}_{\langle P \rangle} \supseteq
\mathbb{R}_{\langle C \rangle}$ to represent the domain of a digital model
of the dynamics of a real plant when evaluated by a verification tool. 
Elements in this domain are marked with an underscore~${\langle C \rangle}$
(see Figure \ref{fig:observersystem}).  Thus, any mathematical operation
within the digital controller is executed in the domain $\mathbb{R}_{\langle
C \rangle}$, and all other calculations in our model will be carried out in
the domain $\mathbb{R}_{\langle P \rangle}$.  Given ${\langle M,F \rangle}$,
we denote as $\mathcal{F}_{\langle M,F \rangle} : \mathbb{R} \rightarrow
\mathbb{R}_{\langle M,F \rangle}$ the corresponding truncation function.  We
further note that the results of mathematical operations in these domains
are also truncated.  This is referred to as roundoff errors.

@@

Since we are interested in synthesizing controllers with observers, in this
section we will look at checking stability, safety and the response time in
the presence of an observer.

%\footnote{A full description of the theory related to observer design is in Appendix \ref{sec:control_theory}.}

@@

We assume that the domain of the observer model is discretized, 
i.e., the domain for both the controller and the observer is $\mathbb{R}_{\langle C \rangle}$.
This means that we have matrices 
${\mat{A}_d}_{\langle C
\rangle},{\mat{B}_d}_{\langle C \rangle},\mat{C}_{\langle C \rangle},
\mat{K}_{\langle C \rangle}, \mat{L}_{\langle C \rangle} \in
\mathbb{R}_{\langle C \rangle}^*$, 
where $*$ represents the appropriate dimension for each matrix.
New problems arise due to the FWL truncation of these matrices. 
These are reflected in the canonical form of the resulting model (see $\mat{A}_e$ and $\mat{B}_e$),
which is expressed by:
%
{\setlength{\belowdisplayskip}{2pt}
\begin{align}
\left [\begin{array}{c}\vec{x}'\\ \vec{x}_e' \end{array}\right]_{k+1}
=& \left [\begin{array}{cc}\mat{A}_{c}+\mat{B}_{e}&\mat{B}_{o}\\ \mat{A}_{e}&\mat{A}_{o}\end{array}\right]
\left [\begin{array}{c}\vec{x}'\\ \vec{x}_e' \end{array}\right]_k
+\mat{B}'\left [\begin{array}{c} \vec{r}_k + \nu_2+\nu_2'\\ \nu_1+\nu_1'+\nu_t\end{array}\right], %\nonumber\\
%\vec{y}_k= &\mat{C}\mat{T}_c\vec{x}_k', 
% + \mat{D}_{d}\vec{u}_k
\label{eq:observer_LTI_FWL_cf}
\end{align}
}
{\setlength{\abovedisplayskip}{2pt}
\setlength{\belowdisplayskip}{2pt}
\begin{align*}
\mat{A}_c=&\left[\begin{array}{cc}
\begin{array}{ccc}k_{p-1}-a_{p-1}&\cdots&k_1-a_1\end{array}&k_0-a_0\\
\mat{I}&\begin{array}{c}0\\ \vdots \\ 0\end{array}
\end{array}\right]&
\mat{A}_o=&\left[\begin{array}{cc}
\begin{array}{c}l_{p-1}-a_{p-1}\\ \vdots\\ l_1-a_1\end{array}&\mat{I}\\
l_0-a_0&\begin{array}{ccc}0&\cdots&0\end{array}
\end{array}\right], 
\end{align*}
}
%
where $\vec{x}'$ is the state of the model in canonical form and
$\vec{x}_e'$ is the error of the observer evaluated in the transformed
domain; further, $\mat{A}_{c}$ the closed-loop matrix of the controller in
canonical form, $\mat{A}_{o}$ the closed-loop matrix of the observer in
canonical form, $\mat{B}_{o}$ the a feedback gain matrix, $\mat{B}'$ the
noise sensitivity matrix, and $\mat{A}_{e}$ and $\mat{B}_{e}$ are the error
caused by the FWL truncation, seen after the transformation to canonical
form.

%and we have introduced the following terms: 
%{\setlength{\abovedisplayskip}{2pt}
%\setlength{\belowdisplayskip}{2pt}
%\begin{align*}
%\mat{A}_{c}=&\mat{T}_c\left(\mat{A}_d-\mat{B}_d\mat{K}_{\langle C \rangle}\right)\mat{T}_c^{-1}&
%\mat{A}_{o}=&\mat{T}_o\left({\mat{A}_d}_{\langle C \rangle}-\mat{L}_{\langle C \rangle}\mat{C}_{\langle C \rangle}\right)\mat{T}_o^{-1}\\
%\mat{B}_{o}=&\mat{T}_c{\mat{B}_d}_{\langle C \rangle}\mat{K}_{\langle C \rangle}\mat{T}_o^{-1}&
%\mat{B}_{e}=&\mat{T}_c(\mat{B}_d-{\mat{B}_d}_{\langle C \rangle})\mat{K}_{\langle C \rangle}\mat{T}_o^{-1}\\
%\mat{A}_{e}=&\mat{T}_o\left(\mat{A}_d-{\mat{A}_d}_{\langle C \rangle}-\mat{L}_{\langle C \rangle}\mat{C}_e\right)\mat{T}_o^{-1}&
%\mat{C}_{e}=&\mat{C}-\mat{C}_{\langle C \rangle}
%\end{align*}
%}
%{\setlength{\abovedisplayskip}{0pt}
%\begin{align*}
%\mat{A}'=&\left [\begin{array}{cc}\mat{A}_{c}+\mat{B}_{e}&\mat{B}_{o}\\ \mat{A}_{e}&\mat{A}_{o}\end{array}\right]&
%\mat{B}'=&\left[\begin{array}{cc}\mat{T}_c{\mat{B}_d}_{\langle C \rangle}&\mat{0}\\\mat{0}&-\mat{T}_o\mat{L}_{\langle C \rangle}\end{array}\right]\\
%\eta_1=&\nu_1+\nu_1'+\nu_t& \eta_2=&\nu_2+\nu_2',
%\end{align*}
%}
%
Moreover, $\nu_1$ and $\nu_2$ are the quantization noises introduced by the
ADC and the DAC, $\nu_t$ is the noise introduced by the delay, and $\nu_1'$
and $\nu_2'$ are quantization noises caused by round-offs in the FWL
operations.

In a real domain $\mathbb{R}$, the characteristic polynomial of the closed
loop dynamics is defined by the characteristic polynomials of~$\mat{A}_c$
and~$\mat{A}_o$.
%
\begin{align*}
P(z)=P_K(z)P_L(z)=\left(z^p+\sum_{i=0}^{p-1}{(a_i-k_i)z^{i}}\right) \left(z^p+\sum_{i=0}^{p-1}{(a_i-l_i)z^{i}}\right)
\end{align*}
%
Jury's criteria uses these coefficients to ensure that no root of the
polynomial has a magnitude greater than one.
%
%\note{we should perhaps provide a qualitative hint on what are the conditions of JC? I'm afraid leaving it to the polynomials here isn't quite complete or clear.}
%
The introduction of the FWL errors causes this to no longer be true. 
We~must therefore find a suitable solution to maintain this independence of
polynomials and use the given coefficients by bounding their error w.r.t. 
the true coefficients of the closed loop matrix.

@@

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Safety of a digitally observed system}\label{sec:cof_safety}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++

A safety specification leads to a requirement on the states of the model. 
Consequently, the controller must ensure that the state never violates the
requirement.  In~this work, the safety property is expressed as:
%
\begin{align}
\phi_\mathit{safety}^n: = & \forall k \in [0, ... , n], \mat{G}\vec{x}_k \leq \vec{h} \text{ and } \underline{y} \leq y_k \leq \overline{y}\nonumber\\
\phi_\mathit{safety}: = & \phi_\mathit{safety}^\infty
\label{eq:safetyspec}
\end{align}
%
where $\mat{G}$ and $\vec{h}$ define a set of inequalities that describe 
a safe set, within the state-space $\mathbb{R}^p$.
%
Note that FWL effects are not explicitly captured here
as they are implicit from Equation \eqref{eq:observer_LTI_FWL_cf}.

@@

%-------------------------------
\subsection{Transient response} 
\label{sec:transientspecification}
%-------------------------------

In many cases the response time of the system, i.e.  how fast it converges
to the set point, is a requirement of great importance.  This is given as a
``settling time''.

\begin{theorem}
%
Given a set of eigenvalues for a discrete time model of a $p^{th}$ order
stable system $\{\lambda_i =e^{\omega_i} \mid i \in [1, ...\,,p],
|\lambda_i| \leq 1 \}$, the settling time of the system is upper-bounded by
%
\begin{equation}
t_s<-log_{\overline{\lambda}}({P_s}), \text{ where }  \overline{\lambda} = \max(|\lambda_i|) \text{ and } \overline{\lambda}<1,\ \forall i \in [1, ...\,,p]
\label{eq:set_time}
\end{equation}
%
where $0\leq P_s \leq 1$ is the desired portion of the target within which
the output must remain after $t_s$.  For the proof see
Appendix~\ref{sec:proofs}
%
\end{theorem}

@@

%The CEGIS-OR loop for this approach is described next.

Throughout the CEGIS-OR loop we make use of a specification $\phi$ that
deals with the abstraction of the model dynamics.  Note that since the model
abstraction gets refined at each iteration, $\phi$ will also be adjusted
accordingly.  Initially, $\phi$ is simply given as the defined $\phi :=
\phi_{stability} \wedge \phi_{input} \wedge \phi_{init}$.

@@

In embedded software programs interacting and/or simulating physical plants
it is of great importance to understand the effects of the chosen numeric
representation for the continuous state-space.  These representations are
almost invariably approximations to the original which are assumed to be
